A few stickleback suffice for the transport of alleles to new lakes

### 2019-09-20 review-responses.tex

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

We are writing to submit a revised version of our manuscript, 
``A few stickleback suffice for the transport of alleles to new lakes''.
Many thanks to the reviewers for thoughtful and helpful comments,
to which we have responded in detail.
We sympathize that additional simulations would be informative and interesting,
but the amount of work that these would entail is beyond our capability at the moment.
We appreciate being explicitly given the option to omit them.
Otherwise, we think that the revisions have helped to clarify and generally improve the manuscript.
We hope you agree.

@@

\vskip 4em

\pagebreak
\setcounter{page}{1}

%%%%%%%%%%%%%%
\reviewersection{AE}

\begin{quote}
    I agree particularly with questions from two of the reviewers asking for
    explanation or additional simulations on the founder population size, with
    reviewer 3's point about clarifying old vs new, and reviewer 1's request
    for more simulations. I do understand that simulations can be time
    consuming, however, and want to make clear additional simulations are not a
    requirement for acceptance (but would be rad if feasible). 
\end{quote}

Additional simulations exploring more of parameter space would be totally awesome, we agree.
However, besides the constraints on our time noted above,
we also note that it's not clear how much additional value it would have to delimit
the region of paramter space on which our results hold:
it would be more satisfying to develop theory that would apply to
larger populations of varying sizes with fluctuating migration rates, for instance.

\begin{point}{\revref}
    My lab and I also spent some time debating what was meant by the explanation at the
    bottom of page 13 -- I think more effort to clarify here would be helpful.
\end{point}

\reply{
    Thanks! We have rewritten this to hopefully be more precise. \revref
}

\begin{point}{}
    Finally, we also wondered why the simulations started at a trait value of 0
    instead of 10 -- it makes comparison of old vs new a bit harder than seems
    necessary. 
\end{point}

\reply{
    We don't have a good reason. Apologies.
}

@@

\plr{IMPORTANT: don't reorder or delete ``points'' below - it messes up the automatic numbering!}

%%%%%%%%%%%%%%
\reviewersection{1}

\begin{quote}
This is a very well done work addressing the fundamental problem of local
adaptation in metapopulations; stickleback are a nice study system but the
findings are relevant for a wide variety of systems. I enjoyed it very much,
and my only wish is that the simulation has been run more than once. This is
not necessary for speed of adaptation - within-simulation replication of lakes
is totally sufficient for this - but for aspects related to genome scanning,
numbers of adaptive alleles in a metapopulation, and haplotype-wise effect
sizes (table 1), since (especially at high M) the number of adaptive alleles is
very few across the whole metapopulation. That said, I do not insist on
rerunning the simulation, since it must be very computationally intensive.
\end{quote}

Thanks for the nice words!
We sympathize, but think that the number of simulations is sufficient
for at least estimating the rough size of these effects -- 
these effects are mostly monotonic in migration rate,
so that for instance we have an estimate of the number of adaptive alleles at $M=1.0$
but also know it is bracketed by the values at $M=0.1$ and $M=10$.
Further simulations would only give us more precise estimates of the mean
and an idea of the standard deviation.

\begin{point}{\revref}
    \textit{``As expected, the genetic basis of the freshwater phenotype seems to simplify
    as migration increases - higher rates of migration allow adaptive alleles of
    larger effect to travel more efficiently through the population, even though
    they are deleterious in the ocean.''}
    Very cool. It gives a very intuitive guideline as to which genetic
    architecture of local adaptation to expect given a certain migration rate.
    There is one neat aspect of this result that is perhaps worth better
    emphasizing: detectable Fst peaks at truly adaptive
    alleles can only be very few. If we had replicate runs we could say how many
    true Fst peaks, on average, we see at M=1 and M=10, Many published studies
    (particularly in marine animals) report dozens if not hundreds of Fst outliers
    throughout the genome; your simulation here suggests that most of them have to
    be false positives of some kind, especially since migration in the sea is
    likely to be high.
\end{point}

\reply{
    We agree with the reviewer that this is an important point,
    and that our results are consistant with the above hypothesis. 
    However, to quantify and assert that previous empirical studies 
    are reporting false-positives would require many 
    more simulations and analysis in many more situations,
    which we believe is outside the realm of this paper.
}

@@

\begin{point}{}
Also, about Fig. S1: I am confused what is ``statistical power'' here :
``likelihood that a SNP will be predicted to have an effect on phenotype when
there is an effect to be detected''. In this model every SNP has at least *some*
effect on phenotype, so what defines ``effect to be detected'' then?
\end{point}

\reply{
    Actually, there are many SNPs that do not affect the phenotype,
    but these were added as a secondary step after simulation,
    so this is an easy mistake to make.
    We've added a reminder about this. \revref
}

\begin{point}{\revref}
    \textit{``Surprisingly, for all cases with the exception of M = 0.01, we found that the
    majority of adaptive allele origins traced back the original generation of
    inhabitants in the lake (Figure 7)''}
    This is really cool and counter-intuitive,
    but maybe it is not too surprising since first-arriving (``captured'') alleles
    are simply much more numerous relative to the alleles that trickle in later,
    *before the adaptation is complete*? Indeed, initial colonization involves
    arrival of 200 individuals in a single generation, which is equivalent to 200/M
    generations of subsequent immigration. Considering that at any appreciable M
    ($>=0.1$) adaptation happens very rapidly (fig 3 and S3), there might simply be
    not enough additional immigrant alleles to make a contribution. I wonder how
    this would work if colonization of a lake was a more gradual process, say 2
    individuals per generation (not all 200 at once)? Would the early-arriving
    alleles still have advantage? (I would bet yes)
\end{point}

\reply{
We agree that this is an important point to explore, and a good observation.
We have appended to the section on migration to more clearly address this \llname{disc_colon}.  
The available code could certainly be used in further studies to investigate this question more in-depth, but again feel that is outside the scope of this paper and have decidecided not to conduct more simulation and analysis.
}

@@

\begin{point}{Title:}
I would make it clear that this is about the ``threespine stickleback'' - there are many different stickleback species.
\end{point}

\reply{
    We believe the first sentence in the abstract makes this clear to the reader.
}

\begin{point}{\revref}
``to millions of independently derived''. Do the references you cite say so? Can you really back up your claim of millions of freshwater populations?
\end{point}

\reply{
    This order of magnitude estimate is roughly based on the length of the coastline
    (roughly 10,000 miles) and the high density of freshwater lakes in many flat, coastal areas.
    We have reworded this to be more conservative, however. \revref
}

@@

\begin{point}{\revref}
P2. ``stickleback in them'' - threespine stickleback in them
\end{point}

\reply{
    Fixed.
}

\begin{point}{p.3 Methods}
From here on, the ms is generally very thinly referenced. In particular, it remains unclear to reader where the many equations come from: are they something authors have derived themselves (I guess not?), or who did that?
\end{point}

\reply{
    We're not sure what is meant here -- the equations in the Methods are almost entirely choices
    we made in setting up the model:
    for instance, the choice of a Gaussian fitness function is almost ubiquitous,
    and the equation for the phenotype in terms of the selection and dominance coefficients
    is just writing down the definition.
    Most of these are options available in SLiM, which we cite.
    We have included a reference for the (simple) defintion of $F_{ST}$ that we use \llname{Fst},
    and added a clarifying comment to the section of the Results and Discussion
    that is most equation-heavy \llname{Gil_cite}.
}

@@

\begin{point}{p.5}
One is left to wonder if it realistic to think that the marine population is as big as 25 freshwater populations. Admittedly, this is discussed to some extent in the discussion, but given the abundance of the threespine stickleback in many seas (there is actually quantitative estimates of this in literature), it seems plausible that the marine population(s) is orders of magnitudes larger than that of typical freshwater population, even if combined. I guess a more quantitative view on this in light of literature would be something worth attempting.
\end{point}

\reply{
    We agree, this would be very interesting, and an informative exercise.
    However, we think it's outside the scope of this paper,
    and hope the reviewer agrees with our discussion of this point at \revref.
}

@@

\begin{point}{\revref}
    Where does the mutation rate $10^{-10}$ come from? Why this particular value, and how are the results influenced by choice of this?
\end{point}

\reply{
    We chose this to reflect our guess at what might be a general consensus for the mutational target size
    of a polygenic trait (although we recognize that opinions differ).
    We've added some more discussion to contextualize our choice to the paper \revref.
    Some more on this: since per-bp mutation rates are likely closer to $10^{-8}$,
    a rate of $10^{-10}$ represents roughly one adaptive target per 100 bases
    across the $10^6$ bases where mutations that affect the phenotype might occur.
    This translates to a total \emph{potential} target size of $10^4$ bp,
    which seems generous but not impossible for a polygenic trait.
}

@@

\begin{point}{}
P4. It is bit unclear here how LD as was dealt with in the simulations. Are all the loci assumed to be unlinked?
\end{point}

\reply{
    All chromsomes are linear with a per-nucleotide recombination rate of $10^{-8}$. This has been clarified here \revref.
}

\begin{point}{\revref}
What happens if generations are overlapping, as is the case in sticklebacks?
\end{point}

\reply{
    We know of no reason why in this case overlapping generation would affect the results.
}

@@

\begin{point}{\revref}
Here the authors say that a different model in SLiM would have been more appropriate to be used, but that they ``do not expect this to influence the results''. This does not sound very good - why not to check this? Why not to rerun the analyses if a more appropriate model is available?
\end{point}

\reply{
In a perfect world, we would re-run and compare. 
However, to re-write this model as non-WF would require a relatively significant amount of change to the code as well as much more computational power to re-run the analysis.
We know of no reason to expect differences between the two implementations --
the main difference being in elegance --
and so have opted to stick with the current model. 
We have altered the wording a bit to better convey this \llname{realistic}.
}

@@

\begin{point}{Fig 1}
The modeled scenario is probably correct over shorter geographic distances. But there is plenty of empirical evidence to show that marine threespine stickleback populations over larger geographic distances are not genetically homogenous. I think it would be appropriate to acknowledge this somewhere in manuscript.
\end{point}

\reply{
We agree, it should be made clear that the oceanic population is not panmictic but rather has one-dimential continuous spatial structure.
We have clarified this here \llname{continuous}.
}

@@

\begin{point}{\revref}
``The efficacy of genomic scans of selection depends on gene flow''. I found this section rather uninformative: all this has been studied in quite detail already in many theoretical (and empirical) papers. This could be shortened. And references to earlier findings would be in place here too - there is no single reference to earlier works here at all.
\end{point}

\reply{
We believe this finding supports existing studies, specifically, how stronger genetic drift with less migration leads to higher background $F_{ST}$.
However, for our study, showing how greater sharing of adaptive alleles provides a shared signal across populations seemed necessary for a study of selection on standing variants.
Thus, we have decided to leave this section as is.
TODO: Looks like we could use some citations, I was unable to find any good ones :(.
}

@@

\begin{point}{\revref}
``Simulation results align with..'' It is nice that the results are referenced back to theory. I was left to think that it might have been better to do this by combining this section with the actual results. Now the structure of the ms appears to be: 1) Results \& Discussion, 2) Discussion. Furthermore, references to theory are given in the first paragraph, but not after this. Readers do not know where all the equations are derived from.
\end{point}

\reply{
    We considered this sort of re-ordering when writing the paper,
    but decided that it would be more confusing to jump back and forth
    to the population genetics heuristics.
    The equations are standard, and we've tried better to flag this at the start of the section.
    \llname{Gil_cite}
}

@@

\begin{point}{\revref}
``we found that standing..'' Is ``found'' right word here? You made an inference, even a guess, that this was so. But you did not prove this?
\end{point}

\reply{
    It is a good point that although we found linkage between opposite-effect alleles,
    this masking might not have a substantial effect.
    We've hedged a bit more in this statement.
}

@@

\begin{point}{\revref}
    ``Population size''. Here it is argued that population sizes of marine and
    freshwater populations are not so different. I found it hard to believe this in
    the light of quantitative data there is. Could it simply be so that marine
    population is actually very much larger, but structured (as there is genetic
    evidence for this in many papers) so that the fraction feeding the freshwater
    colonizations is to some degree isolated from ones feeding the freshwaters in a
    particular area? This would in effect make relevant portion of the marine
    population smaller, without need to postulate that the total marine population
    equals in size to freshwater populations, and that there are millions of lakes
    inhabited by three spine sticklebacks. Note also that in introduction it was
    claimed millions of freshwater populations, here you say ``if not millions'' -
    inconsistent. 
\end{point}

\reply{
    We've made the two sections consistent,
    and still heartily agree that quantifying this would be an interesting and useful endeavor,
    but beyond the scope of the current paper.
    The proposed substructure in the ocean doesn't seem very plausible to us,
    although there is little enough data on oceanic demography that this may well be correct.
    Still, is our argument that they face similar constraints on breeding ground not convincing?
}

@@

\begin{quote}
    The paper is well written and I like this work and don't see any
    fundamental flaws in the way the simulations were set up or in how the results
    are interpreted. Of course, one always needs to make certain assumptions when
    setting up simulations (such as on effect sizes of loci etc.), but I think that
    the authors explore the parameter space appropriately. 
\end{quote}

Thanks! We're glad you liked it.

\begin{quote}
    In what follows, I provide suggestions on how to further improve the work. Most
    of these points should be understood as recommendations. A side-note:
    Please make sure to add line numbers to your next submission. It is
    difficult for reviewers to refer to specific text passages without line
    numbers being present.
\end{quote}

Sorry about that; we're submitting with line numbers this time.

@@

\begin{point}{}
Founder effects followed by population growth: I wonder how founder effects during the colonization of the 'new lakes', followed by population growth, would influence the results? It seems that this was never explored, as each new lake had a constant population size of 200 individuals (rather low). I don't expect there to be a qualitative change in the overall results, but the relative contribution of de novo vs. standing variants to adaptation might well be influenced by a more colonization scenario. If the authors decided not to run additional simulations on this issue, it might at least be worth speculating on how founder effects followed by population growth might influence some of their results (such as the results shown in Fig. 7).
\end{point}

\reply{
Good point: we agree this would be one of the first additional parameters to vary.
We have discussed this further \llname{disc_colon},
but as we explained above, haven't done more simulations to expand coverage of parameter space.
}

@@

\begin{point}{}
If possible, it would be interesting to check what type of mutations - in terms of effect size and in terms of dominance - fall under the 'captured', 'marine' and 'de novo' categories. I could well imagine, for example, that mutations that can stay in the marine populations relatively long are more likely to be recessive. Is this true?
\end{point}

\reply{
    We thought about this also, and did this comparison in some earlier simulations,
    but any differential representation was not obvious
    -- there was no large effect --
    and that we'd need many more simulations to detect anything.
    So: this seems plausible, but speculatively,
    it doesn't seem to be a major driver of genetic architecture,
    but rather a minor adjustment.
}

@@

\begin{point}{}
I'm a bit confused by how the adaptive loci along the chromosome were modeled. I don't fully understand what you mean by mutations occurring at a rate $10^{-10}$ per locus per generations in ten regions of $10^5$ loci each (Page 4, 'Genetic architecture of the trait'). Why are the adaptive loci themselves organized in distinct regions? Why isn't there simply a given number of adaptive loci that are evenly spread across the chromosome?
\end{point}

\reply{
    We agree this was confusing and have edited this section to hopefully make it more clear \revreffull{2}{8}. 
    The motivation for 10 regions of effect was to roughly model the distribution of QTL seen in experiments,
    and to allow us to contrast closely linked variants (within the same region)
    with more distant ones.
}

@@

\begin{point}{Results \& Discussion section}
    Sometimes, it is not fully clear whether you talk about the 'old' lakes or the 'new' lakes. This is often unclear, for example, in the section ``Rapid local adaptation at intermediate gene flow rates''. Could you make this clear throughout the manuscript?
\end{point}

\reply{
Good point, we have clarified a few ambiguous spots, e.g., \revref.
}

@@

\begin{point}{Figure 2}
    Are these the results after 50k generations?
\end{point}

\reply{
    Actually, at the time of introduction (so 100,000 generations), thank you for the clarification, this has been noted in the figure caption. 
}

\begin{point}{}
    In the legend of Figure 4, it might be good to re-iterate what the number on top of the panels means exactly. Say that these are actual numbers of migrants per population and what the population size is. What is meant by 'M' is said more clearly in the legend to Figure 6, yet, it would be good to here also state (again) how large the populations are. M can only be understood in context of the actual population size. Otherwise, you would need to provide 'm'.
\end{point}

\reply{
    We have revised the figure caption and added the label $M$ to the header.
}

### 2019-09-20 Stickleback_Paper.tex
fixed plot
peter edits

The canonical model for the genetics of adaptation has long been the sequential fixation of new mutations 
\citep{smith_haigh_1974, Endler_1977, Orr_2005}.
While it has proved to be a useful baseline for understanding the genetic variation we see in species today, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures 
\citep{Lai2152, Schrider2017}. 
In particular, empirical studies have increasingly identified the need to more deeply incorporate standing genetic variation into adaptation dynamics for metapopulations inhabiting an array of habitats 
\citep{Hermisson_2008, Barrett2008}. 
Populations experiencing diverse selective pressures while still exhibiting significant gene flow often result in more complex genomic signals that are still not fully understood 
\citep{Charlesworth_1997, Charlesworth_2003, Nosil_2009, Flaxman_2013, Samuk_2017}. 
Concurrently, a growing number of empirical studies have identified instances of convergent evolution using standing genetic variation 
\citep{jones_genomic_2012, roesti_genome_2012, feulner_genomics_2015, Schrider2017, Barrett2008, Nelson_2018, Nelson_2019, Bassham_2018}. \revpoint{3}{10}
However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of standing genetic variation, particularly during colonization and adaptation to new environments
\citep{Nelson_2018, Pritchard_2010, yeaman2011genetic, Schrider2017}. 
It is similarly unclear whether variation in these evolutionary processes can determine the genetic architecture of evolving traits via standing genetic variation. 
Theoretical work is mostly limited to one or two loci, or relies on approximations of uncertain validity
\citep[e.g.,][]{slatkin1975selection,petry1983effect,barton1986barrier,yeaman2011genetic,ralph2015standing}.
Therefore, detailed simulation models of specific systems can 
provide an important complement to empirical studies in the lab and field,
by helping us understand precisely how standing genetic variation might fuel local adaptation,
and what genomic signals we can expect to see.

The ancestral marine form of threespine stickleback fish (\textit{Gasterosteus aculeatus}) \revpoint{2}{1}
has given rise to hundreds of thousands if not millions of independently derived freshwater populations \revpoint{2}{3} \llabel{hundreds-thousands-or-more}
in recently de-glaciated regions around the Northern Hemisphere 
\citep{Bell_1994, Thompson_1997, Cresko_2007, Hunt_2008} \revpoint{3}{9}
This model organism has provided some of the earliest data showing the heterogeneous nature of divergence across genomes \revpoint{2}{4}
and the much more extensive use of standing genetic variation than once thought 
\citep{schluter2009genetics, DeFaveri_2011, DeFaveri_2013, Roesti_2014, Nelson_2018, Bassham_2018, Nelson_2019, Hohenlohe_2010, Terekhanova_2014, Marques_2016}.  
While geographic isolation often prevents direct migration between freshwater populations, threespine stickleback in them \revpoint{2}{5}
frequently evolve similar phenotypes \citep{Cresko_2004, Colosimo_2004}. 
The most recent evolutionary genomic studies on stickleback document that while the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years 
\citep{Nelson_2018}, independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations \citep{Terekhanova_2014, Lescak2015evolution, Bassham_2018}. 
For example, in 1964 the Great Alaskan Earthquake caused an uplift of many islands and coastal regions throughout the Gulf of Alaska. 
Studies of stickleback populations on uplifted Middleton Island showed that newly created freshwater ponds were invaded by the surrounding marine population of stickleback which evolved the freshwater syndrome of phenotypes in less than 50 years 
\citep{Lescak2015evolution, Bassham_2018}. 
Amazingly, the portions of the genomes of these populations that showed increased divergence from the oceanic population
mirrored those previously found to differ between ocean and freshwater populations that
that have been geographically separated for thousands of years 
\citep{Hohenlohe_2010, Bassham_2018}.

@@

To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} to implement forward-time simulation of populations of individuals with explicitly represented genomes in which selection acted upon a single continuous quantitative trait. 
The details and equations of the model were motivated by current understanding of threespine stickleback evolutionary, history and demography, as well as population genetics theory \citep{Gillespie_1998}. \llabel{clarify_equations}
The following model included divergent selection in the two habitats which had substantial spatial structure.
Certain aspects of the model remain simplistic due to computational constraints;
in particular, total population sizes are much smaller than in reality,
although we may capture a good picture of local dynamics.

\paragraph{Habitat and geography}
Our simulations include two habitat types -- marine and freshwater -- defined by the nature of their selective pressures, each with 5,000 diploid individuals. 
The arrangement of these habitats, depicted in Figure~\ref{fig:Geo}, roughly models a set of freshwater habitats along a stretch of coastline. 
The marine habitat is a continuous, one-dimensional range of length 25 units, while the freshwater habitat is divided into 25 discrete subpopulations (which we call ``lakes''), each connected to the marine habitat at regularly spaced intervals (positions $i - 1/2$ for $1 \le i \le 25$).

@@

Each individual carries two linear chromosomes, each of size $10^8$ loci,
with recombination events occurring at a intralocus rate of $10^{-8}$ every generation. \revpoint{2}{9}
%Mutations that can affect the trait under selection can occur at rate $10^{-10}$ per locus per generation \revpoint{2}{8}
%in ten regions of $10^5$ loci each, spread evenly along the chromosome. 
Mutations that can affect the trait under selection can occur at rate $10^{-8}$ with a target size of $10$ Kb, (spread evenly among $1$ Mb) across each individual genome. \revpoint{2}{8}
With 5000 (diploid) individuals per population, this equates to $\approx 1$ mutation, per population, per generation, which impact individual trait value. \llabel{mu_exp}
Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability). 
Effect sizes for these mutations are chosen randomly from an exponential distribution with mean $1/2$, either positive or negative with equal probability. 
Individual trait values ($x_\text{ind}$) are determined additively from the diploid genotypes. 
Concretely, an individual that is heterozygous and homozygous for mutations at sets of loci $H$ and $D$ respectively has trait value $x_\text{ind} = \sum_{i \in H} h_i s_i + \sum_{j \in D} s_j$, where $h_i$ and $s_i$ are the dominance coefficient and the effect size of the mutation at locus $i$. Subsequent mutations at the same locus replace the previous allele.

@@

We use SLiM to simulate a Wright--Fisher population with non-overlapping generations \revpoint{2}{10}
and a fixed population size of 5,000 diploid individuals in each habitat. 
Each generation, the two parents of each new offspring are chosen proportional to their fitness 
(unlike actual stickleback, all individuals are hermaphroditic), and the contributing genomes are produced by Poisson recombination with an average of one crossover per chromosome per generation ($10^{-8}$ per locus per generation). 
Since the total population across \emph{all} 25 lakes is fixed at 5000, and the Wright--Fisher model assumes unrealistic global population regulation, we normalize the fitnesses of each individual so that approximately 200 offspring are generated in each lake, each generation. 
(A more realistic implementation would use SLiM's ``non-Wright-Fisher'' model type, \revpoint{2}{11}
which was not available when we wrote the simulations; but we do not expect the difference to affect results.) \llabel{realistic}
To do this, we divide fitness values of each freshwater individual by the mean fitness in their lake, so that the mean fitnesses of all lakes are equal before selection happens.

As depicted in Figure~\ref{fig:Geo} the marine population is continuous (unidimensional) while lakes are discrete. \llabel{continuous}
Dispersal occurs both locally along the coastline within the marine habitat, as well as between the marine habitat and the lakes,
with a lake--ocean migration rate denoted $m$. 
No migration occurs directly between the freshwater populations. 
All individual dispersal events can be thought of as occurring at the juvenile stage in the life cycle of the simulation. 
Each new individual in each habitat has parents from the other habitat with probability $m$ (in which case we call it a ``migrant''), and parents from the same habitat with probability $1-m$.
The first parent of each non-migrant individual in the freshwater habitat is chosen from the freshwater habitat proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness. The offspring then lives in the same lake as the parents. Parents for each non-migrant marine individual are chosen similarly:
first, a single parent is chosen proportionally to fitness in the marine habitat, and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$. Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$, where $f(x_j)$ is the fitness of individual $j$ and $d_{ij}$ is the distance between the two locations. 
Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.5, and reflected to stay within the habitat range.
Parents for each freshwater migrant are chosen in the same way as for non-migrant marine individuals, and are assigned to the lake nearest to the position of the first marine parent. Similarly, parents for each marine migrant are both chosen from the same lake as before, and the offspring is given a spatial location in the marine habitat at the location of the parent's lake.

@@

\begin{figure}
    \begin{center}
        \includegraphics{Final_Plots/Pheno_Time.pdf}
          \caption{ 
        Mean individual trait values in the marine habitat (red line),
                the original lakes (tan lines; average thicker),
                and the new lakes (black lines; average thicker),
                across the course of two simulations, with migration rates of
                \textbf{(A)} $M=0.01$ and
                \textbf{(B)} $M=0.1$ 
                migrants per lake per generation, respectively.
                Optimal trait values in the two habitats are at $\pm 10$.
                    Analogous plots for other migration rates
                    are shown in Figure \ref{fig:phenotype_time_supp}.
        }
          \label{fig:phenotype_time}
    \end{center}
\end{figure}

>>>>

The canonical model for the genetics of adaptation has long been the sequential fixation of new mutations 
\citep{smith_haigh_1974, Endler_1977, Orr_2005}.
While it has proved to be a useful baseline for understanding the genetic variation we see in species today, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures 
\citep{Lai2152, Schrider2017}. 
In particular, empirical studies have increasingly identified the need to more deeply incorporate standing genetic variation into adaptation dynamics for metapopulations inhabiting an array of habitats 
\citep{Hermisson_2008, Barrett2008}. 
Populations experiencing diverse selective pressures while still exhibiting significant gene flow often result in more complex genomic signals that are still not fully understood 
\citep{Charlesworth_1997, Charlesworth_2003, Nosil_2009, Flaxman_2013, Samuk_2017}. 
Concurrently, a growing number of empirical studies have identified instances of convergent evolution using standing genetic variation 
\citep{jones_genomic_2012, roesti_genome_2012, feulner_genomics_2015, Schrider2017, Barrett2008, Nelson_2018, Nelson_2019, Bassham_2018}. \revpoint{3}{10}
However, it is still not clear how variation in evolutionary processes -- such as gene flow, recombination, selection and mutation --  can promote the maintenance and re-use of standing genetic variation, particularly during colonization and adaptation to new environments
\citep{Nelson_2018, Pritchard_2010, yeaman2011genetic, Schrider2017}. 
It is similarly unclear whether variation in these evolutionary processes can determine the genetic architecture of evolving traits via standing genetic variation. 
Theoretical work is mostly limited to one or two loci, or relies on approximations of uncertain validity
\citep[e.g.,][]{slatkin1975selection,petry1983effect,barton1986barrier,yeaman2011genetic,ralph2015standing}.
Therefore, detailed simulation models of specific systems can 
provide an important complement to empirical studies in the lab and field,
by helping us understand precisely how standing genetic variation might fuel local adaptation,
and what genomic signals we can expect to see.

The ancestral marine form of threespine stickleback fish (\textit{Gasterosteus aculeatus}) \revpoint{2}{1}
has given rise to hundreds of thousands to millions of independently derived freshwater populations \revpoint{2}{3}
in recently de-glaciated regions around the Northern Hemisphere 
\citep{Bell_1994, Thompson_1997, Cresko_2007, Hunt_2008} \revpoint{3}{9}
This model organism has provided some of the earliest data showing the heterogeneous nature of divergence across genomes \revpoint{2}{4}
and the much more extensive use of standing genetic variation than once thought 
\citep{schluter2009genetics, DeFaveri_2011, DeFaveri_2013, Roesti_2014, Nelson_2018, Bassham_2018, Nelson_2019, Hohenlohe_2010, Terekhanova_2014, Marques_2016}.  
While geographic isolation often prevents direct migration between freshwater populations, threespine stickleback in them \revpoint{2}{5}
frequently evolve similar phenotypes \citep{Cresko_2004, Colosimo_2004}. 
The most recent evolutionary genomic studies on stickleback document that while the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years 
\citep{Nelson_2018}, independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations \citep{Terekhanova_2014, Lescak2015evolution, Bassham_2018}. 
For example, in 1964 the Great Alaskan Earthquake caused an uplift of many islands and coastal regions throughout the Gulf of Alaska. 
Studies of stickleback populations on uplifted Middleton Island showed that newly created freshwater ponds were invaded by the surrounding marine population of stickleback which evolved the freshwater syndrome of phenotypes in less than 50 years 
\citep{Lescak2015evolution, Bassham_2018}. 
Amazingly, the portions of the genomes of these populations that showed increased divergence from the oceanic population
mirrored those previously found to differ between ocean and freshwater populations that
that have been geographically separated for thousands of years 
\citep{Hohenlohe_2010, Bassham_2018}.

@@

To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} to implement forward-time simulation of populations of individuals with explicitly represented genomes in which selection acted upon a single continuous quantitative trait. 
The details of the model were motivated by current understanding of threespine stickleback evolutionary, history and demography.
The following model included divergent selection in the two habitats which had substantial spatial structure.
Certain aspects of the model remain simplistic due to computational constraints;
in particular, total population sizes are much smaller than in reality,
although we may capture a good picture of local dynamics.

\paragraph{Habitat and geography}
Our simulations include two habitat types -- marine and freshwater -- defined by the nature of their selective pressures, each with 5,000 diploid individuals. 
The arrangement of these habitats, depicted in \autoref{fig:Geo}, roughly models a set of freshwater habitats along a stretch of coastline. 
The marine habitat is a continuous, one-dimensional range of length 25 units, while the freshwater habitat is divided into 25 discrete subpopulations (which we call ``lakes''), each connected to the marine habitat at regularly spaced intervals (positions $i - 1/2$ for $1 \le i \le 25$).

@@

Individuals are diploid, with a linear chromosome $10^8$ bp long,
with recombination events occurring at a rate of $10^{-8}$ per bp per generation. \revpoint{2}{9}
Mutations that can affect the trait under selection occur at rate $10^{-10}$ per locus per generation
in ten regions of $10^5$ loci each, spread evenly along the chromosome.
This translates to a total adaptive target size of $10^{-4}$ per individual per generation, \revpoint{2}{8}
or $\approx 1$ trait-affecting mutation, per population of 5,000 diploids, per generation.
This architecture was chosen to model a moderate number of clusters of trait-affecting genes,
each of which having about 1Kb of potentially adaptive target
(since mutation rates are closer to $10^{-8}$, a mutation rate of $10^{-10}$
is roughly equivalent to having potentially adaptive sites at one of every 100 bases).
Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability). 
Effect sizes for these mutations are chosen randomly from an exponential distribution with mean $1/2$, either positive or negative with equal probability. 
Individual trait values ($x_\text{ind}$) are determined additively from the diploid genotypes. 
Concretely, an individual that is heterozygous and homozygous for mutations at sets of loci $H$ and $D$ respectively has trait value $x_\text{ind} = \sum_{i \in H} h_i s_i + \sum_{j \in D} s_j$, where $h_i$ and $s_i$ are the dominance coefficient and the effect size of the mutation at locus $i$. Subsequent mutations at the same locus replace the previous allele.

@@

We use SLiM to simulate a Wright--Fisher population with non-overlapping generations \revpoint{2}{10}
and a fixed population size of 5,000 diploid individuals in each habitat. 
Each generation, the two parents of each new offspring are chosen proportional to their fitness 
(unlike actual stickleback, all individuals are hermaphroditic), and the contributing genomes are produced by Poisson recombination with an average of one crossover per chromosome per generation ($10^{-8}$ per locus per generation). 
Since the total population across \emph{all} 25 lakes is fixed at 5000, and the Wright--Fisher model assumes unrealistic global population regulation, we normalize the fitnesses of each individual so that approximately 200 offspring are generated in each lake, each generation. 
(A simpler implementation would use SLiM's ``non-Wright-Fisher'' model type, \revpoint{2}{11}
which was not available when we wrote the simulations; but we do not expect the difference to affect results.) \llabel{realistic}
To do this, we divide fitness values of each freshwater individual by the mean fitness in their lake, so that the mean fitnesses of all lakes are equal before selection happens.

As depicted in \autoref{fig:Geo} the marine population is continuous (one-dimensional) while lakes are discrete. \llabel{continuous}
Dispersal occurs both locally along the coastline within the marine habitat, as well as between the marine habitat and the lakes,
with a lake--ocean migration rate denoted $m$. 
No migration occurs directly between the freshwater populations. 
All individual dispersal events can be thought of as occurring at the juvenile stage in the life cycle of the simulation. 
Each new individual in each habitat has parents from the other habitat with probability $m$ (in which case we call it a ``migrant''), and parents from the same habitat with probability $1-m$.
The first parent of each non-migrant individual in the freshwater habitat is chosen from the freshwater habitat proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness. The offspring then lives in the same lake as the parents. Parents for each non-migrant marine individual are chosen similarly:
first, a single parent is chosen proportionally to fitness in the marine habitat, and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$. Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$, where $f(x_j)$ is the fitness of individual $j$ and $d_{ij}$ is the distance between the two locations. 
Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.5, and reflected to stay within the habitat range.
Parents for each freshwater migrant are chosen in the same way as for non-migrant marine individuals, and are assigned to the lake nearest to the position of the first marine parent. Similarly, parents for each marine migrant are both chosen from the same lake as before, and the offspring is given a spatial location in the marine habitat at the location of the parent's lake.

@@

\begin{figure}
    \begin{center}
        \includegraphics{Final_Plots/Pheno_Time.pdf}
          \caption{ 
        Mean individual trait values in the marine habitat (red line),
                the original lakes (tan lines; average thicker),
                and the new lakes (black lines; average thicker),
                across the course of two simulations, with migration rates of
                \textbf{(A)} $M=0.01$ and
                \textbf{(B)} $M=0.1$ 
                migrants per lake per generation, respectively.
                Optimal trait values in the two habitats are at $\pm 10$.
                    Analogous plots for other migration rates
                    are shown in \autoref{fig:phenotype_time_supp}.
        }
          \label{fig:phenotype_time}
    \end{center}
\end{figure}
